### 2016-01-07 mcmc.tex
gibbs implementation note

As we mentioned above, scientists often like to report the ``best sample'';
  that is, the sample with the highest posterior pdf value.
This is not usually a good idea.
For one, MCMC is a sampling algorithm, not an optimization algorithm;
  there are no guarantees that it will find the optimum of the posterior pdf in reasonable time.%
  \note{The sampler samples the pdf fairly.
    The tallest peak in the posterior is not guaranteed to be---%
    and in general is not---%
    also the peak with the greatest posterior ``mass''.
    That is, the samples in a fair sampling will not necessarily come
    predominantly from the \emph{tallest} peak in the posterior pdf.
    They will come from the peak with the greatest total integral
    of the likelihood over the prior.
    These issues might sound ``academic'',
    but when the number of parameters gets large (greater than, say, 10),
    many normal intuitions one might have (if any) about what is reasonably likely in a pdf
    are regularly violated.}
For two, the closeness of the best sample to the posterior mode is a strong function of sample size,
  and with very bad scaling.%
  \note{If you have a $K$-point sampling in a $D$-dimensional parameter space,
    there is some best sample $\pars_k^{\best}$.  Now imagine that you want to make
    that best sample 10 times better---that is, you want to find a point $\pars^{\better}$
    that is ten times closer to the true optimum.  How many more samples do you need to take?
    In the limit that $K$ is large, on average, you have to take a factor of
    $10^D$ more samples!
    Any reasonable optimizer is far, far better than this (in terms of compute time).}
For three, if you just want an optimum, use an optimizer!%
  \note{HOGG, something?}
That is, we \emph{strongly advise against} reporting,
  as ``the'' measurement or ``the result'',
  the parameter value or values for the best sample.
Giving some posterior samples is a good idea (see below).
If you feel drawn to give the best sample,
  instead give the parameter values found by optimizing the posterior PDF,
  starting at the best sample as an initialization
  (and sacrifice your probabilism\note{something}).
This optimal-posterior parameter value is called the
  ``maximum \foreign{a posteriori}'' or ``MAP'' value for the parameters;
  it is like a maximum-likelihood value but regularized by the prior.
It is an estimator with some good (and some bad) properties, but it is not the point of MCMC,
  and therefore outside the scope of this \documentname.

@@

A trivial kind of Gibbs sampling was mentioned above in the
discussion of M--H MCMC:
Even when running vanilla M--H MCMC, it is often useful to update only
one parameter at a time; that is, do only axis-aligned moves in the
$\pars$-space, cycling through axes as you go from step to step.
This helps with tuning and diagnosis, but it also permits a good
implementation to capitalize on simpler calculations for the simpler
parameter directions.

%% HOGG:  Note here that this requires writing the f(\pars) code to capitalize on the axis-aligned moves!!

>>>>

As we mentioned above, scientists often like to report the ``best sample'';
  that is, the sample with the highest posterior pdf value.
This is not usually a good idea.
For one, MCMC is a sampling algorithm, not an optimization algorithm;
  there are no guarantees that it will find the optimum of the posterior pdf in reasonable time.%
  \note{The sampler samples the pdf fairly.
    The tallest peak in the posterior is not guaranteed to be---%
    and in general is not---%
    also the peak with the greatest posterior ``mass''.
    That is, the samples in a fair sampling will not necessarily come
    predominantly from the \emph{tallest} peak in the posterior pdf.
    They will come from the peak with the greatest total integral
    of the likelihood over the prior.
    These issues might sound ``academic'',
    but when the number of parameters gets large (greater than, say, 10),
    many normal intuitions one might have (if any) about what is reasonably likely in a pdf
    are regularly violated.}
For two, the closeness of the best sample to the posterior mode is a strong function of sample size,
  and with very bad scaling.%
  \note{If you have a $K$-point sampling in a $D$-dimensional parameter space,
    there is some best sample $\pars_k^{\best}$.  Now imagine that you want to make
    that best sample 10 times better---that is, you want to find a point $\pars^{\better}$
    that is ten times closer to the true optimum.  How many more samples do you need to take?
    In the limit that $K$ is large, on average, you have to take a factor of
    $10^D$ more samples!
    Any reasonable optimizer is far, far better than this (in terms of compute time).}
For three, if you just want an optimum, use an optimizer!
That is, we \emph{strongly advise against} reporting,
  as ``the'' measurement or ``the result'',
  the parameter value or values for the best sample.
Giving some posterior samples is a good idea (see below).
If you feel drawn to give the best sample,
  instead give the parameter values found by optimizing the posterior PDF,
  starting at the best sample as an initialization
  (and sacrifice your probabilism\note{something}).
This optimal-posterior parameter value is called the
  ``maximum \foreign{a posteriori}'' or ``MAP'' value for the parameters;
  it is like a maximum-likelihood value but regularized by the prior.
It is an estimator with some good (and some bad) properties, but it is not the point of MCMC,
  and therefore outside the scope of this \documentname.

@@

A trivial kind of Gibbs sampling was mentioned above in the
discussion of M--H MCMC:
Even when running vanilla M--H MCMC, it is often useful to update only
one parameter at a time; that is, do only axis-aligned moves in the
$\pars$-space, cycling through axes as you go from step to step.
This helps with tuning and diagnosis, but it also permits a good
implementation to capitalize on simpler calculations for the simpler
parameter directions.\note{%
  The point is that sometimes there are some parameters that only
  affect a very small number of data points, or parameters that change
  the model or the likelihood function and prior values in a very
  simple-to-compute way.  If such simple parameters exist, it makes
  sense to write your likelihood and prior code to capitalize on these
  simplicities.  This is not always trivial, because it may involve
  caching parts of the calculation and so on.  Details would be beyond
  our scope.}
